### 2020-04-08 5_evaluation.tex
More tweaks to reduce page count.

METG is useful in part because it summarizes the overhead of each programming system in a single number. This makes it
possible to evaluate METG at different node counts (shown in
Figure~\ref{fig:metg-compute}) to see how it is impacted by changes in
communication topology and latency. A key finding is that
the systems with the smallest METG on one node have roughly an order
of magnitude higher METG at 256 nodes---increased communication latencies require significantly larger tasks to
achieve the same level of efficiency, so apparent differences in
runtime overhead at small node counts can matter much less or not at
all at larger node counts.

@@

PaRSEC, StarPU and Regent
rely on runtime analysis that can suffer from
scalability bottlenecks if every node must consider the tasks
executing on all other nodes. PaRSEC DTD and StarPU mitigate this
partially by allowing the user to omit tasks not directly
dependent on those executed by a given node. However, this DAG
trimming approach requires dynamic checks that scale with
the number of nodes and thus limit scalability~\cite{PARSEC_DTD}. Compile-time analysis can
partially or fully mitigate this overhead. PaRSEC PTG improves over DTD 
by performing DAG expansion at compile time~\cite{PARSEC_DTD}, but
retains dynamic checks that limit scalability. PaRSEC shard includes additional manual optimizations over DTD, completely eliminating these dynamic checks. Regent
uses a compile-time optimization to improve
scalability~\cite{ControlReplication17}; the increase in METG beyond 16 nodes is due to a known bug in Realm barrier migration which the Realm Task Bench implementation is able to manually work around.

@@

The ratio in METG between 0 and 3 dependencies per task ranges from
$1.01\times$ to $250\times$ (median $2.9\times$). The difference is most pronounced in systems that
perform runtime work inline. For example, MPI achieves an METG of 390
ns with 0 dependencies, but this rises to 4.6 \textmu{}s with 3 dependencies,
a $12\times$ increase. This is unsurprising as in the case
of 0 dependencies, no \lstinline[language=C++]{MPI_Isend} calls are
issued at all, so MPI has nothing to do aside from executing kernels. Clearly, choosing a representative dependence
pattern is important when estimating the performance of a workload or
class of workloads.

@@

To determine the cost of scheduling tasks on GPUs, Figure~\ref{fig:cuda-efficiency} compares MPI
and MPI+CUDA on the Piz Daint supercomputer. The CUDA
compute kernel achieves a peak performance of $4.759 \times 10^{12}$
FLOP/s, which is very close to the officially reported number
$4.761 \times 10^{12}$. The CPU achieves $5.726 \times 10^{11}$
FLOP/s. Note that the kernels perform different numbers of
operations as the GPU requires more work to reach peak
performance. The x-axis in Figure~\ref{fig:cuda-efficiency} is
normalized to keep FLOPs constant for a given problem
size.

Our MPI+CUDA implementation uses an offload model in which data is copied to and from the GPU on each step. We test
two configurations: \lstinline{w1} uses 1 task per GPU, whereas
\lstinline{w4} overdecomposes by $4\times$, using 4 MPI
ranks per GPU to push work to the GPU in parallel.
\lstinline{w4} achieves higher FLOP/s but
drops more rapidly at smaller problem sizes, due to the overhead of
scheduling $4\times$ as many CUDA kernel launches. In either case, the
GPU requires more work to achieve high performance, and the overhead
of copying data dominates at small task
granularities, where the CPU achieves higher
performance. While Figure~\ref{fig:cuda-efficiency} is not couched in
terms of METG (as peak performance on CPU and GPU are very
different), the conclusion here is similar to
Section~\ref{subsec:scalability}: the cost of sending data and
tasks to the GPU imposes a higher task granularity to achieve
the same efficiency than running on the CPU, reducing the advantage at small task granularities of
very lightweight mechanisms such as those in MPI.

>>>>

METG summarizes system overheads in a single number. This makes it possible to
evaluate how communication topology and latency impact METG at
different node counts, as shown in Figure~\ref{fig:metg-compute}.
We find that
systems with the smallest METG on one node have roughly an order
of magnitude higher METG at 256 nodes. Increased communication latencies require significantly larger tasks to
achieve the same level of efficiency, so apparent differences in
overhead at small node counts can matter much less or not at
all at larger node counts.

@@

PaRSEC, StarPU and Regent
rely on runtime analysis that can suffer from
scalability bottlenecks if every node must consider the tasks
executing on all other nodes. PaRSEC DTD and StarPU mitigate this
partially by allowing the user to omit tasks not directly
dependent on those executed by a given node. However, this approach still requires checks that scale with
node count and thus limit scalability~\cite{PARSEC_DTD}. Compile-time analysis can
partially or fully mitigate this overhead. PaRSEC PTG improves over DTD 
by performing DAG expansion at compile time~\cite{PARSEC_DTD}, but
retains dynamic checks that limit scalability. PaRSEC shard includes additional manual optimizations over DTD, completely eliminating these dynamic checks. Regent
uses a compile-time optimization to improve
scalability~\cite{ControlReplication17}; the increase in METG beyond 16 nodes is due to a known bug in Realm barrier migration which the Realm Task Bench implementation is able to manually work around.

@@

The ratio in METG between 0 and 3 dependencies per task ranges from
$1.01\times$ to $250\times$ (median $2.9\times$). The difference is most pronounced in systems that
perform runtime work inline. For example, MPI achieves an METG of 390
ns with 0 dependencies, but this rises to 4.6 \textmu{}s with 3 dependencies,
a $12\times$ increase. This is unsurprising, as with
0 dependencies no \lstinline[language=C++]{MPI_Isend} calls are
issued at all. Clearly, choosing a representative dependence
pattern is important when estimating the performance of a workload or
class of workloads.

@@

To determine the cost of scheduling tasks on GPUs, Figure~\ref{fig:cuda-efficiency} compares MPI
and MPI+CUDA on Piz Daint~\cite{PizDaint}. The CUDA
compute kernel achieves $4.759 \times 10^{12}$
FLOP/s, which is very close to the officially reported number
$4.761 \times 10^{12}$. The CPU achieves $5.726 \times 10^{11}$
FLOP/s. Note that the kernels perform different numbers of
operations as the GPU requires more work to reach peak
performance. The x-axis in Figure~\ref{fig:cuda-efficiency} is
normalized to keep FLOPs constant for a given problem
size.

Our MPI+CUDA implementation uses an offload model with data copied to and from the GPU on each step. We test
two configurations: \lstinline{w1} uses 1 task per GPU, whereas
\lstinline{w4} overdecomposes, using 4 MPI
ranks per GPU to push work to the GPU in parallel.
\lstinline{w4} achieves higher FLOP/s but
drops more rapidly at small problem sizes, due to the overhead of
scheduling $4\times$ as many CUDA kernel launches. Either way,
GPUs require more work to achieve high performance, and the overhead
of copying data dominates at small task
granularities, where CPUs achieve higher
performance. While Figure~\ref{fig:cuda-efficiency} is not couched in
terms of METG (as peak performance on CPU and GPU are very
different), the conclusion here is similar to
Section~\ref{subsec:scalability}: the cost of sending data and
tasks to the GPU imposes a floor on task granularity relative to the CPU, reducing the advantage at small task granularities of
very lightweight mechanisms such as those in MPI.

### 2020-04-08 8_case_study.tex
More tweaks to reduce page count.

Realm and Chapel both use DMA subsystems that are optimized for large
copies. Early Task Bench experiments revealed high METG values that
were diagnosed by the Realm/Chapel developers as overhead due to the
cost of scheduling small copies. Subsequent improvements in Realm and
Chapel improved small copy overheads (and thus METG) by over an order
of magnitude in the case of Realm, and by $2\times$ in the case of
Chapel. These improvements affect any application where fine-grained
data movement is needed, which is particularly relevant in strong scaling
regimes when running on large numbers of nodes.

>>>>

Realm and Chapel both use DMA subsystems optimized for large
copies. Early Task Bench experiments revealed high METG values,
diagnosed by the Realm/Chapel developers as overhead due to the
cost of scheduling small copies. Subsequent improvements in Realm and
Chapel improved small copy overheads (and thus METG) by over an order
of magnitude in the case of Realm, and by $2\times$ in the case of
Chapel. These improvements affect any application where fine-grained
data movement is needed, which is particularly relevant in strong scaling
regimes when running on large numbers of nodes.
